---
title: "Migration"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message=FALSE, echo=TRUE)
```

## Migration Flows Sankey

As well as supporting the generation of parameterised reports, reprodducble workflwos also support the automated generation of (templated) code that implements interactive charts.

For example, inspired by Oli Hawkins ([Visualising migration between the countries of the UK](http://olihawkins.com/2017/03/1); [demo](http://olihawkins.com/visualisation/8)),  we can generate interactive Sankey plots using the `googleVis` or `rCharts` packages, .


```{r}
#The RCharts package thros a wobbly if we don't load knitr in explicitly
library(knitr)
library(readr)

#Data from ONS: https://www.ons.gov.uk/peoplepopulationandcommunity/populationandmigration/migrationwithintheuk/datasets/matricesofinternalmigrationmovesbetweenlocalauthoritiesandregionsincludingthecountriesofwalesscotlandandnorthernireland
regionsquarematrix2015 = read_csv("./laandregionsquarematrices2015/regionsquarematrix2015.csv", skip = 8)

#The data has thousand separator commas - so remove them and convert to numeric
#There is probably a more iddiomatic way of doing this using tidyr...
regionsquarematrix2015 = cbind(regionsquarematrix2015[1:2], sapply(regionsquarematrix2015[3:ncol(regionsquarematrix2015)], function(x)  as.numeric(gsub(",", "", x)) ) )

head(regionsquarematrix2015)
```

The Sankey diagram generates expect the data to be provided as edge lists (*from*, *to*, *value*).

```{r}
library(tidyr)
#Melt the data(wide to long) so we have from/to/value flows
rr=regionsquarematrix2015 %>% gather(to, value, 3:ncol(.))

#The googleVis sankey diagram hates cycles - so set unique labels for from/to
#Not sure if we can use separate labels/identifier columns so we can label things the same?
rr$to=paste0(rr$to,'_')
```

Generate a Sankey diagram using `googleVis`:

```{r, results='asis'}
library(googleVis)
options(gvis.plot.tag='chart')
#Generate the Sankey diagram HTML
s=gvisSankey(rr[!is.na(rr$value),c('Region','to','value')], from="Region", to="to", weight="value")
#And render it
plot(s)
```

Generate a Sankey diagram using `rCharts`:

```{r, results='asis'}
#Based on http://bl.ocks.org/timelyportfolio/6085852
#There is also a particle flow enhancement demoed at https://bl.ocks.org/micahstubbs/6a366e759f029599678e293521d7e26c
library(rCharts)

colnames(rr) = c("Name","source","target","value")

sankeyPlot2 <- rCharts$new()
sankeyPlot2$setLib('http://timelyportfolio.github.io/rCharts_d3_sankey/')
sankeyPlot2$set(
  data = rr,
  nodeWidth = 15,
  nodePadding = 10,
  layout = 32,
  width = 750,
  height = 500
)
sankeyPlot2$show('iframesrc', cdn = TRUE)
```

A wide range of interactive chart types can be generated in this way. The [*htmlwidgets for R*](http://www.htmlwidgets.org/) project represents the latest iteration in the production of interactive Javascript widgets for use in RMarkdown documents and Shiny applications.