---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message=FALSE, echo=TRUE)
```

## Migration Flows Sankey

As well as supporting the generation of parameterised reports, reproducible workflows also support the automated generation of (templated) code that implements interactive charts.

For example, inspired by Oli Hawkins ([Visualising migration between the countries of the UK](http://olihawkins.com/2017/03/1) [[demo](http://olihawkins.com/visualisation/8)]),  we can generate interactive Sankey plots using the `googleVis` or `rCharts` packages, .


```{r}
#The RCharts package thros a wobbly if we don't load knitr in explicitly
library(knitr)
library(readr)

#Data from ONS: https://www.ons.gov.uk/peoplepopulationandcommunity/populationandmigration/migrationwithintheuk/datasets/matricesofinternalmigrationmovesbetweenlocalauthoritiesandregionsincludingthecountriesofwalesscotlandandnorthernireland
regionsquarematrix2015 = read_csv("./laandregionsquarematrices2015/regionsquarematrix2015.csv", skip = 8)

#The data has thousand separator commas - so remove them and convert to numeric
#There is probably a more idiomatic way of doing this using tidyr...
regionsquarematrix2015 = cbind(regionsquarematrix2015[1:2], sapply(regionsquarematrix2015[3:ncol(regionsquarematrix2015)], function(x)  as.numeric(gsub(",", "", x)) ) )

head(regionsquarematrix2015)
```

The Sankey diagram generators seem to expect the data to be provided as edge lists (*from*, *to*, *value*).

```{r}
library(tidyr)
#Melt the data(wide to long) so we have from/to/value flows
rr=regionsquarematrix2015 %>% gather(to, value, 3:ncol(.))

#The Sankey diagram generators dislike cycles - so set unique labels for from/to
#Not sure if we can use separate labels/identifier columns so we can label things the same?
rr$to=paste0(rr$to,'_')
```

### `googleVis`

Generate a Sankey diagram using `googleVis`:

```{r, results='asis'}
#For use in Rmd/knitr, set the block parameter: results='asis'
library(googleVis)
options(gvis.plot.tag='chart')
#Generate the Sankey diagram HTML
s=gvisSankey(rr[!is.na(rr$value),c('Region','to','value')], from="Region", to="to", weight="value")
#And render it
plot(s)
```

According to the [Google Sankey diagram documentation](https://developers.google.com/chart/interactive/docs/gallery/sankey), node labels, as well as the color of nodes and edges, can be controlled; see [Sankey diagrams with googleVis](http://www.magesblog.com/2014/03/sankey-diagrams-with-googlevis.html) or [this StackOverflow answer](http://stackoverflow.com/a/32111596/454773) for an example of how to pass the parameters in.

### `rCharts`

Generate a Sankey diagram using `rCharts`:

```{r, results='asis'}
#Based on http://bl.ocks.org/timelyportfolio/6085852
#There is also a particle flow enhancement demoed at https://bl.ocks.org/micahstubbs/6a366e759f029599678e293521d7e26c
library(rCharts)

colnames(rr) = c("Name","source","target","value")

sankeyPlot2 <- rCharts$new()
sankeyPlot2$setLib('http://timelyportfolio.github.io/rCharts_d3_sankey/')
sankeyPlot2$set(
  data = rr,
  nodeWidth = 15,
  nodePadding = 10,
  layout = 32,
  width = 750,
  height = 500
)
sankeyPlot2$show('iframesrc', cdn = TRUE)
#Note that at the time of writing, the rCharts_d3_sankey bakes in the http protocol for loading three
#resources that breaks if the output HTML page is served as https.
```

Some control over colouring can be introduced by extending the template, as demonstrated in [this Stack Overflow answer](http://stackoverflow.com/a/28500397/454773) (following [this original explanation](http://stackoverflow.com/a/25416927/454773) from @timelyportfolio, the author of the rCharts Sankey package.)

A wide range of interactive chart types can be generated in this way. The [*htmlwidgets for R*](http://www.htmlwidgets.org/) project represents the latest iteration in the production of interactive Javascript widgets for use in RMarkdown documents and Shiny applications.

### `sankeyD3`

This seems to be the most recent attempt at an R/Sankey diagram library, again using D3.js.

```{r, results='asis'}
#devtools::install_github("fbreitwieser/sankeyD3")
library(sankeyD3)

#BROKEN: I think the edgelist (Links) *requires* that nodes are identified as integers 0..N-1
sankeyNetwork(Links = rr, Nodes = rr, Source = "source",
            Target = "target", Value = "value", NodeID = "Name",
              fontSize = 12, nodeWidth = 30)
```